\documentclass[11pt,a4paper]{article}
\usepackage{shared/parscover}

% --- Packages ---
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{fancyhdr}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{listings}
\usepackage{microtype}
\usepackage{natbib}
\usepackage{doi}
\usepackage{url}
\usepackage{xspace}
\usepackage{multirow}

\geometry{margin=1in}
\pgfplotsset{compat=1.18}

% --- Macros ---
\newcommand{\Pars}{\textsc{Pars}\xspace}
\newcommand{\parsfl}{\textsc{ParsFL}\xspace}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\EE}{\mathbb{E}}
\newcommand{\PP}{\mathbb{P}}
\newcommand{\NN}{\mathcal{N}}

\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}

% --- Header ---
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small Cyrus Pahlavi, Pars Team}
\fancyhead[R]{\small Privacy-Preserving Federated Learning}
\fancyfoot[C]{\thepage}

\title{Privacy-Preserving Federated Learning\\for Diaspora AI Applications\\[0.5em]
\large Secure Aggregation and Differential Privacy on the Pars Network}

\author{
  Cyrus Pahlavi, Pars Team\\
  \texttt{research@pars.id}
}

\date{February 2026}

\begin{document}
\parscoverpage


\begin{abstract}
We present \parsfl, a federated learning framework designed for privacy-preserving collaborative model training across the geographically dispersed Pars Network. \parsfl addresses the unique challenges of diaspora-scale federated learning: extreme data heterogeneity arising from diverse linguistic and cultural contexts, communication constraints imposed by censorship-resistant networking, and the heightened privacy requirements of users in adversarial jurisdictions. Our framework combines threshold secret-sharing-based secure aggregation with R\'enyi differential privacy accounting, enabling Byzantine-resilient training without revealing individual contributions. We introduce \emph{adaptive communication compression} that reduces bandwidth by 94\% compared to naive gradient transmission while preserving convergence guarantees. Theoretical analysis establishes convergence bounds under non-IID data distributions with Byzantine participants, and empirical evaluation on Persian NLP and community health tasks demonstrates that \parsfl achieves 96.2\% of centralized training accuracy while providing $(\varepsilon=2, \delta=10^{-6})$-differential privacy and tolerating up to 20\% Byzantine nodes.
\end{abstract}

\section{Introduction}

Federated learning (FL) enables collaborative model training across distributed devices without centralizing raw data~\citep{mcmahan2017communication,kairouz2021advances}. For the Pars Network, FL is a critical capability: community AI applications---content moderation, search personalization, language models---benefit from training on diverse community data, but centralizing this data would create unacceptable privacy risks for users in adversarial jurisdictions.

The diaspora setting introduces challenges beyond standard FL scenarios:

\begin{enumerate}[label=(\arabic*)]
    \item \textbf{Extreme heterogeneity:} Data distributions vary dramatically across geographic regions, languages (Farsi, Dari, Tajiki, Finglish), and usage patterns.
    \item \textbf{Communication constraints:} Many participants connect through censorship-resistant channels with limited bandwidth and high latency.
    \item \textbf{Adversarial participants:} State-sponsored actors may join the network to poison models or extract information about other participants.
    \item \textbf{Variable availability:} Users in censored regions have intermittent connectivity, leading to high dropout rates during training rounds.
    \item \textbf{Privacy under coercion:} Users may face device seizure; even encrypted local data could be compelled to be decrypted.
\end{enumerate}

\subsection{Contributions}

\begin{enumerate}[label=(\roman*)]
    \item \textbf{Secure Aggregation Protocol:} A threshold secret-sharing scheme that provides information-theoretic privacy against coalitions of up to $t < n/3$ colluding parties, with $O(n \log n)$ communication complexity per round.
    \item \textbf{Differential Privacy Integration:} R\'enyi differential privacy (RDP) accounting with adaptive clipping that provides formal $(\varepsilon, \delta)$-DP guarantees per participant while minimizing utility loss.
    \item \textbf{Byzantine Resilience:} A coordinate-wise trimmed mean aggregation rule that tolerates up to 20\% Byzantine participants without significantly degrading convergence.
    \item \textbf{Adaptive Compression:} A communication-efficient protocol combining top-$k$ sparsification with quantization, reducing per-round communication by 94\% while maintaining convergence.
    \item \textbf{Convergence Theory:} Tight convergence bounds for non-IID, Byzantine, compressed, and differentially private federated learning.
\end{enumerate}

\section{Background}

\subsection{Federated Averaging}

The FedAvg algorithm~\citep{mcmahan2017communication} is the canonical FL method. Given $n$ clients with local datasets $\{D_1, \ldots, D_n\}$, FedAvg minimizes:

\begin{equation}
    \min_{w \in \RR^d} F(w) = \sum_{i=1}^{n} \frac{|D_i|}{|D|} F_i(w), \quad F_i(w) = \frac{1}{|D_i|} \sum_{(x,y) \in D_i} \ell(w; x, y)
\end{equation}

Each round $t$, the server broadcasts $w^t$, each client performs $\tau$ local SGD steps, and the server averages the updates:

\begin{equation}
    w^{t+1} = w^t + \frac{1}{n} \sum_{i=1}^{n} \Delta_i^t, \quad \Delta_i^t = w_i^{t,\tau} - w^t
\end{equation}

\subsection{Secure Aggregation}

Secure aggregation~\citep{bonawitz2017practical} enables the server to compute the sum $\sum_i \Delta_i$ without learning individual $\Delta_i$. The approach uses pairwise masking: each pair $(i, j)$ agrees on a random mask $r_{ij}$, and client $i$ submits $\hat{\Delta}_i = \Delta_i + \sum_{j \neq i} s_{ij} \cdot r_{ij}$ where $s_{ij} = +1$ if $i < j$ and $-1$ otherwise. The masks cancel in summation.

\subsection{Differential Privacy}

\begin{definition}[Differential Privacy~\citep{dwork2006calibrating}]
A randomized mechanism $\mathcal{M}$ is $(\varepsilon, \delta)$-differentially private if for all adjacent datasets $D, D'$ and all measurable sets $S$:
\begin{equation}
    \PP[\mathcal{M}(D) \in S] \leq e^\varepsilon \cdot \PP[\mathcal{M}(D') \in S] + \delta
\end{equation}
\end{definition}

\begin{definition}[R\'enyi Differential Privacy~\citep{mironov2017renyi}]
$\mathcal{M}$ satisfies $(\alpha, \varepsilon)$-RDP if for all adjacent $D, D'$:
\begin{equation}
    D_\alpha(\mathcal{M}(D) \| \mathcal{M}(D')) = \frac{1}{\alpha - 1} \log \EE\left[\left(\frac{\PP[\mathcal{M}(D) = o]}{\PP[\mathcal{M}(D') = o]}\right)^\alpha\right] \leq \varepsilon
\end{equation}
\end{definition}

\subsection{Byzantine-Resilient Aggregation}

Byzantine-resilient aggregation rules replace the simple average with robust estimators. Krum~\citep{blanchard2017machine}, trimmed mean~\citep{yin2018byzantine}, and median~\citep{yin2018byzantine} provide different resilience-utility tradeoffs.

\section{System Model}

\subsection{Network Architecture}

\parsfl operates on a star topology overlaid on the Pars mesh network. The aggregation server is replicated across $m = 5$ geographically distributed nodes using threshold multi-party computation (MPC). This eliminates the single server as a point of failure or privacy compromise.

\begin{definition}[Distributed Aggregation Committee]
The aggregation committee $\mathcal{C} = \{c_1, \ldots, c_m\}$ consists of $m$ nodes that collectively perform secure aggregation using $(t_c, m)$-threshold secret sharing, where $t_c = \lceil m/2 \rceil + 1$. The committee is selected via verifiable random function (VRF) from the set of qualifying relay nodes each epoch.
\end{definition}

\subsection{Threat Model}

\begin{itemize}
    \item \textbf{Semi-honest aggregators:} Up to $t_c - 1$ aggregation committee members may be corrupted and will attempt to learn individual updates but follow the protocol.
    \item \textbf{Byzantine clients:} Up to $f < n/5$ clients may send arbitrary updates to degrade model quality or extract information.
    \item \textbf{External observer:} A network-level adversary can observe encrypted traffic patterns but cannot decrypt contents.
    \item \textbf{Post-hoc compromise:} A client's device may be compromised after participation; DP ensures limited information leakage about the local dataset.
\end{itemize}

\section{Secure Aggregation Protocol}

\subsection{Overview}

Our secure aggregation protocol proceeds in four phases:

\begin{algorithm}[H]
\caption{\parsfl Secure Aggregation}
\label{alg:secagg}
\begin{algorithmic}[1]
\Require Clients $\{1, \ldots, n\}$ with updates $\{\Delta_1, \ldots, \Delta_n\}$, committee $\mathcal{C}$
\Ensure Aggregated update $\bar{\Delta} = \frac{1}{n}\sum_i \Delta_i$ known only to $\mathcal{C}$ collectively
\Statex \textbf{Phase 1: Key Agreement}
\For{each client pair $(i, j)$}
    \State $(pk_i, sk_i) \leftarrow \mathsf{KeyGen}()$
    \State Exchange $pk_i, pk_j$ via committee relay
    \State $r_{ij} \leftarrow \mathsf{PRG}(\mathsf{DH}(sk_i, pk_j))$
\EndFor
\Statex \textbf{Phase 2: Masked Submission}
\For{each client $i$}
    \State $\hat{\Delta}_i \leftarrow \Delta_i + \sum_{j \neq i} s_{ij} \cdot r_{ij} + z_i$ \Comment{$z_i$: DP noise share}
    \State Secret-share $\hat{\Delta}_i$ among committee: $[\hat{\Delta}_i]_1, \ldots, [\hat{\Delta}_i]_m$
    \State Send $[\hat{\Delta}_i]_k$ to committee member $c_k$
\EndFor
\Statex \textbf{Phase 3: Committee Aggregation}
\For{each committee member $c_k$}
    \State $[S]_k \leftarrow \sum_{i=1}^{n} [\hat{\Delta}_i]_k$ \Comment{Homomorphic sum of shares}
\EndFor
\Statex \textbf{Phase 4: Reconstruction}
\State Committee reconstructs $S = \sum_i \hat{\Delta}_i$ from $\{[S]_k\}_{k=1}^{m}$
\State Masks cancel: $S = \sum_i \Delta_i + \sum_i z_i = \sum_i \Delta_i + z$
\State $\bar{\Delta} \leftarrow S / n$
\end{algorithmic}
\end{algorithm}

\subsection{Dropout Handling}

Client dropout during training is common in the diaspora setting. We handle dropouts using the technique from~\citep{bonawitz2017practical}: each client also secret-shares their mask seed $\mathsf{DH}(sk_i, pk_j)$ among the committee during Phase 1. If client $i$ drops out before Phase 2 completion, the committee can reconstruct $i$'s masks and subtract them from the partial sum.

\begin{algorithm}[H]
\caption{Dropout Recovery}
\label{alg:dropout}
\begin{algorithmic}[1]
\Require Active clients $A \subseteq [n]$, dropped clients $D = [n] \setminus A$
\Ensure Correct aggregation over active clients
\For{$i \in D$}
    \State Committee reconstructs mask seeds for client $i$
    \For{$j \in A$}
        \State Compute mask $r_{ij}$ from reconstructed seed
        \State Subtract $s_{ij} \cdot r_{ij}$ from partial sum
    \EndFor
\EndFor
\State Compute $\bar{\Delta} = \frac{1}{|A|} \sum_{i \in A} \Delta_i + z$
\end{algorithmic}
\end{algorithm}

\subsection{Communication Complexity}

\begin{theorem}[Communication Complexity]
\parsfl secure aggregation requires $O(n \cdot d / k + n \cdot m)$ total communication per round, where $d$ is the model dimension, $k$ is the sparsification factor, and $m$ is the committee size.
\end{theorem}

\begin{proof}
Each client sends $d/k$ compressed gradient coordinates to each of $m$ committee members (after sparsification), plus $O(m)$ overhead for mask seed shares. Total per-client communication is $O(d/k \cdot m + m)$. Summing over $n$ clients gives $O(n \cdot d \cdot m / k + n \cdot m)$. Since $m$ is constant (5), this simplifies to $O(n \cdot d / k + n)$.
\end{proof}

\section{Differential Privacy}

\subsection{Per-Round Privacy}

Each client clips their update and adds calibrated Gaussian noise before submission:

\begin{algorithm}[H]
\caption{DP Update Preparation}
\label{alg:dp}
\begin{algorithmic}[1]
\Require Client update $\Delta_i$, clipping bound $C$, noise multiplier $\sigma$
\Ensure DP-protected update $\tilde{\Delta}_i$
\State $\Delta_i^{\text{clip}} \leftarrow \Delta_i \cdot \min\left(1, \frac{C}{\|\Delta_i\|_2}\right)$ \Comment{L2 clipping}
\State $z_i \sim \NN(0, \sigma^2 C^2 I_d / n^2)$ \Comment{Gaussian noise (distributed)}
\State $\tilde{\Delta}_i \leftarrow \Delta_i^{\text{clip}} + z_i$
\State \Return $\tilde{\Delta}_i$
\end{algorithmic}
\end{algorithm}

The total noise added to the aggregate is $z = \sum_i z_i \sim \NN(0, \sigma^2 C^2 I_d / n)$, which is equivalent to centralized DP-SGD~\citep{abadi2016deep} noise.

\subsection{Adaptive Clipping}

Fixed clipping bounds $C$ can significantly degrade utility when gradient norms vary across rounds. We use the adaptive clipping method of~\citep{andrew2021differentially}:

\begin{algorithm}[H]
\caption{Adaptive Clipping Bound}
\label{alg:adaptive-clip}
\begin{algorithmic}[1]
\Require Current bound $C^t$, target quantile $\gamma$ (default 0.5), learning rate $\eta_C$
\Ensure Updated bound $C^{t+1}$
\State Each client $i$ reports $b_i = \mathbf{1}[\|\Delta_i\|_2 \leq C^t]$ (with DP noise)
\State Server computes $\hat{\gamma} = \frac{1}{n} \sum_i b_i$
\State $C^{t+1} \leftarrow C^t \cdot \exp(-\eta_C (\hat{\gamma} - \gamma))$
\end{algorithmic}
\end{algorithm}

\subsection{R\'enyi DP Accounting}

We track privacy loss using R\'enyi differential privacy (RDP) for tighter composition:

\begin{theorem}[Per-Round RDP Guarantee]
Each round of \parsfl with Gaussian noise multiplier $\sigma$, subsampling rate $q = k/n$, and clipping bound $C$ satisfies $(\alpha, \varepsilon(\alpha))$-RDP where:
\begin{equation}
    \varepsilon(\alpha) \leq \frac{1}{\alpha - 1} \log\left(1 + q^2 \binom{\alpha}{2} \frac{2}{\sigma^2} + \sum_{j=3}^{\alpha} q^j \binom{\alpha}{j} \frac{j(j-1)}{2\sigma^2}\right)
\end{equation}
\end{theorem}

After $T$ rounds, RDP composes linearly: the total RDP cost is $T \cdot \varepsilon(\alpha)$. We convert to $(\varepsilon, \delta)$-DP using the optimal conversion~\citep{balle2020hypothesis}:

\begin{equation}
    \varepsilon = \min_{\alpha > 1} \left\{ T \cdot \varepsilon(\alpha) + \frac{\log(1/\delta)}{\alpha - 1} \right\}
\end{equation}

\subsection{Privacy Budget Allocation}

\begin{table}[H]
\centering
\caption{Privacy Budget for Different Application Scenarios}
\label{tab:privacy-budget}
\begin{tabular}{lcccc}
\toprule
\textbf{Application} & \textbf{Rounds} & $\sigma$ & $\varepsilon$ & $\delta$ \\
\midrule
Content moderation & 500 & 1.0 & 2.0 & $10^{-6}$ \\
Language model fine-tune & 1{,}000 & 1.2 & 3.5 & $10^{-6}$ \\
Search personalization & 200 & 0.8 & 1.2 & $10^{-6}$ \\
Community health & 300 & 1.5 & 1.8 & $10^{-7}$ \\
\bottomrule
\end{tabular}
\end{table}

\section{Byzantine Resilience}

\subsection{Coordinate-Wise Trimmed Mean}

We use coordinate-wise trimmed mean as our robust aggregation rule:

\begin{algorithm}[H]
\caption{Byzantine-Resilient Aggregation}
\label{alg:byzantine}
\begin{algorithmic}[1]
\Require Updates $\{\Delta_1, \ldots, \Delta_n\}$, Byzantine fraction bound $\beta$
\Ensure Robust aggregate $\bar{\Delta}$
\State $b \leftarrow \lceil \beta \cdot n \rceil$ \Comment{Number of values to trim}
\For{each coordinate $j \in \{1, \ldots, d\}$}
    \State Sort $\{\Delta_{1,j}, \ldots, \Delta_{n,j}\}$ in ascending order
    \State Remove $b$ smallest and $b$ largest values
    \State $\bar{\Delta}_j \leftarrow \frac{1}{n - 2b} \sum_{i=b+1}^{n-b} \Delta_{(i),j}$
\EndFor
\State \Return $\bar{\Delta}$
\end{algorithmic}
\end{algorithm}

\subsection{Compatibility with Secure Aggregation}

Trimmed mean requires access to individual coordinate values, which conflicts with secure aggregation's goal of hiding individual updates. We resolve this through a two-stage protocol:

\begin{enumerate}
    \item \textbf{Norm reporting:} Each client reports $\|\Delta_i\|_2$ (with DP noise) to the committee. Updates with anomalous norms (beyond 3 standard deviations) are flagged.
    \item \textbf{Bucketed aggregation:} The coordinate space is divided into $B$ buckets. For each bucket, the committee performs secure aggregation separately, allowing outlier detection at the bucket level without revealing individual coordinates.
    \item \textbf{Fallback:} If more than $\beta$ fraction of clients are flagged, the round is aborted and clients are asked to resubmit with reduced local training steps.
\end{enumerate}

\subsection{Byzantine Resilience Guarantee}

\begin{theorem}[Byzantine Resilience]
Under Assumption~\ref{ass:bounded} (bounded gradient variance) and with at most $f < n/5$ Byzantine clients, \parsfl with coordinate-wise trimmed mean converges to a neighborhood of the optimal solution:
\begin{equation}
    \EE[F(w^T)] - F^* \leq O\left(\frac{\sigma^2}{\sqrt{nT}} + \frac{f^2}{n^2} \cdot G^2\right)
\end{equation}
where $\sigma^2$ is the gradient variance, $G$ is the gradient bound, and $T$ is the number of rounds.
\end{theorem}

\section{Adaptive Communication Compression}

\subsection{Compression Pipeline}

Communication is the primary bottleneck in diaspora FL. We apply a three-stage compression pipeline:

\begin{algorithm}[H]
\caption{Adaptive Compression}
\label{alg:compress}
\begin{algorithmic}[1]
\Require Update $\Delta_i \in \RR^d$, sparsification ratio $s$, quantization bits $b$
\Ensure Compressed update $\hat{\Delta}_i$, error feedback $e_i$
\State $\Delta_i' \leftarrow \Delta_i + e_i$ \Comment{Error feedback from previous round}
\Statex \textbf{Stage 1: Top-$k$ Sparsification}
\State $k \leftarrow \lfloor s \cdot d \rfloor$
\State $S \leftarrow \text{indices of top-}k \text{ coordinates of } |\Delta_i'|$
\State $\Delta_i^{\text{sparse}} \leftarrow \Delta_i'|_S$ \Comment{Keep only top-$k$}
\Statex \textbf{Stage 2: Stochastic Quantization}
\For{each $j \in S$}
    \State $\hat{\Delta}_{i,j}^{\text{sparse}} \leftarrow \mathsf{StochasticQuantize}(\Delta_{i,j}^{\text{sparse}}, b)$
\EndFor
\Statex \textbf{Stage 3: Entropy Coding}
\State $\hat{\Delta}_i \leftarrow \mathsf{ArithmeticEncode}(S, \hat{\Delta}_i^{\text{sparse}})$
\Statex \textbf{Error Feedback}
\State $e_i \leftarrow \Delta_i' - \mathsf{Decompress}(\hat{\Delta}_i)$
\State \Return $(\hat{\Delta}_i, e_i)$
\end{algorithmic}
\end{algorithm}

\subsection{Adaptive Sparsification Ratio}

The sparsification ratio adapts to available bandwidth:

\begin{equation}
    s^t = \min\left(s_{\max}, \frac{B_i \cdot T_{\text{round}}}{d \cdot b \cdot \log_2(d)}\right)
\end{equation}

where $B_i$ is client $i$'s available bandwidth, $T_{\text{round}}$ is the round deadline, $d$ is the model dimension, and $b$ is the quantization bit width.

\subsection{Compression Ratio}

\begin{table}[H]
\centering
\caption{Communication Compression Results}
\label{tab:compression}
\begin{tabular}{lccc}
\toprule
\textbf{Method} & \textbf{Bits/param} & \textbf{Compression} & \textbf{Accuracy Drop} \\
\midrule
No compression (fp32) & 32.0 & 1$\times$ & 0.0\% \\
fp16 & 16.0 & 2$\times$ & 0.1\% \\
Top-1\% sparsification & 0.32 & 100$\times$ & 1.2\% \\
Top-1\% + 8-bit quant. & 0.08 & 400$\times$ & 1.8\% \\
\parsfl adaptive (ours) & 0.05 & 640$\times$ & 2.1\% \\
\parsfl + error feedback & 0.05 & 640$\times$ & 0.8\% \\
\bottomrule
\end{tabular}
\end{table}

\section{Convergence Analysis}

\subsection{Assumptions}

\begin{enumerate}[label=\textbf{A\arabic*}]
    \item \label{ass:smooth} \textbf{Smoothness:} Each $F_i$ is $L$-smooth: $\|\nabla F_i(w) - \nabla F_i(v)\| \leq L \|w - v\|$.
    \item \label{ass:bounded} \textbf{Bounded variance:} $\EE[\|\nabla F_i(w) - \nabla F(w)\|^2] \leq \sigma_G^2$ (gradient heterogeneity) and $\EE[\|g_i - \nabla F_i(w)\|^2] \leq \sigma_L^2$ (stochastic noise).
    \item \label{ass:compress} \textbf{Compression:} The compression operator $\mathcal{C}$ satisfies $\EE[\|\mathcal{C}(x) - x\|^2] \leq (1 - \delta_c) \|x\|^2$ for some $\delta_c > 0$.
\end{enumerate}

\subsection{Main Convergence Result}

\begin{theorem}[Convergence of \parsfl]
\label{thm:convergence}
Under Assumptions A1--A3, with $n$ clients ($f$ Byzantine), compression parameter $\delta_c$, DP noise variance $\sigma_{\text{DP}}^2$, learning rate $\eta = O(1/\sqrt{T})$, and $\tau$ local steps per round, \parsfl satisfies:
\begin{equation}
    \frac{1}{T} \sum_{t=0}^{T-1} \EE[\|\nabla F(w^t)\|^2] \leq O\left(\frac{F_0 - F^*}{\eta T} + \eta L \sigma_L^2 + \eta^2 L^2 \tau^2 \sigma_G^2 + \frac{\sigma_{\text{DP}}^2}{n} + \frac{f^2 G^2}{n^2 \delta_c}\right)
\end{equation}
where $F_0 = F(w^0)$.
\end{theorem}

\begin{proof}[Proof Sketch]
The proof combines three analyses: (1)~standard FedAvg convergence under heterogeneity~\citep{li2020federated}, extended with (2)~the compression error analysis from~\citep{stich2018sparsified} using error feedback, and (3)~the Byzantine resilience bound from~\citep{yin2018byzantine} with the DP noise treated as an additional variance term. The full proof is in Appendix~\ref{app:proof}.
\end{proof}

\begin{corollary}[Convergence Rate]
With optimal learning rate $\eta = O(n^{-1/2} T^{-1/2})$ and $T = \Theta(n/\varepsilon_{\text{opt}}^2)$ rounds:
\begin{equation}
    \frac{1}{T} \sum_{t=0}^{T-1} \EE[\|\nabla F(w^t)\|^2] = O\left(\frac{1}{\sqrt{nT}} + \frac{\sigma_{\text{DP}}^2}{n}\right)
\end{equation}
achieving near-linear speedup in the number of clients when the DP noise term is subdominant.
\end{corollary}

\section{Implementation}

\subsection{System Architecture}

\parsfl is implemented as a modular framework in Python (client) and Rust (aggregation server):

\begin{table}[H]
\centering
\caption{System Components}
\label{tab:system}
\begin{tabular}{lll}
\toprule
\textbf{Component} & \textbf{Language} & \textbf{Description} \\
\midrule
Client SDK & Python (PyTorch) & Local training, compression, DP \\
Aggregation server & Rust & Secure aggregation, Byzantine filter \\
Communication layer & Rust (QUIC) & Encrypted transport \\
Coordinator & Go & Round management, client selection \\
Privacy accountant & Python (Opacus) & RDP tracking, budget management \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Client-Side Training}

\begin{algorithm}[H]
\caption{Client Training Loop}
\label{alg:client}
\begin{algorithmic}[1]
\Require Global model $w^t$, local dataset $D_i$, local epochs $\tau$, clipping bound $C$
\Ensure Compressed, clipped, noised update $\hat{\Delta}_i$
\State $w_i \leftarrow w^t$
\For{$e = 1$ to $\tau$}
    \For{batch $B \subset D_i$}
        \State $g \leftarrow \nabla \ell(w_i; B)$
        \State Per-sample gradient clipping: $g \leftarrow g \cdot \min(1, C / \|g\|_2)$
        \State $w_i \leftarrow w_i - \eta_{\text{local}} \cdot g$
    \EndFor
\EndFor
\State $\Delta_i \leftarrow w_i - w^t$
\State $\tilde{\Delta}_i \leftarrow \mathsf{ClipAndNoise}(\Delta_i, C, \sigma)$ \Comment{Algorithm~\ref{alg:dp}}
\State $\hat{\Delta}_i \leftarrow \mathsf{Compress}(\tilde{\Delta}_i)$ \Comment{Algorithm~\ref{alg:compress}}
\State \Return $\hat{\Delta}_i$
\end{algorithmic}
\end{algorithm}

\section{Evaluation}

\subsection{Experimental Setup}

We evaluate \parsfl on four tasks relevant to the Pars Network:

\begin{table}[H]
\centering
\caption{Evaluation Tasks}
\label{tab:tasks}
\begin{tabular}{llccc}
\toprule
\textbf{Task} & \textbf{Model} & \textbf{Params} & \textbf{Clients} & \textbf{Data/Client} \\
\midrule
Persian sentiment & BERT-base & 110M & 200 & 500--5K \\
Content moderation & ResNet-18 & 11M & 500 & 200--2K \\
NER (Persian) & BiLSTM-CRF & 8M & 100 & 1K--10K \\
Language model & GPT-2 Small & 124M & 50 & 10K--100K \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Data Heterogeneity}

We simulate diaspora-scale heterogeneity using Dirichlet distribution with concentration parameter $\alpha_{\text{Dir}}$:

\begin{table}[H]
\centering
\caption{Effect of Data Heterogeneity on Convergence}
\label{tab:heterogeneity}
\begin{tabular}{lcccc}
\toprule
$\alpha_{\text{Dir}}$ & \textbf{IID Level} & \textbf{FedAvg} & \textbf{FedProx} & \textbf{\parsfl} \\
\midrule
$\infty$ (IID) & Perfect & 91.2 & 91.0 & 90.8 \\
1.0 & Moderate & 87.4 & 88.1 & 88.6 \\
0.1 & High & 78.3 & 82.7 & 85.1 \\
0.01 & Extreme & 62.1 & 71.4 & 79.3 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Privacy-Utility Tradeoff}

\begin{table}[H]
\centering
\caption{Privacy-Utility Tradeoff (Persian Sentiment, 200 clients)}
\label{tab:privacy-utility}
\begin{tabular}{lccc}
\toprule
$\varepsilon$ & \textbf{Accuracy} & \textbf{vs.\ Central} & \textbf{Noise $\sigma$} \\
\midrule
$\infty$ (no DP) & 89.4\% & 97.2\% & 0 \\
8.0 & 88.7\% & 96.4\% & 0.5 \\
4.0 & 87.9\% & 95.6\% & 0.8 \\
2.0 & 86.3\% & 93.8\% & 1.0 \\
1.0 & 83.1\% & 90.3\% & 1.5 \\
0.5 & 78.4\% & 85.2\% & 2.0 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Byzantine Resilience}

\begin{table}[H]
\centering
\caption{Byzantine Resilience (Persian Sentiment, 200 clients)}
\label{tab:byzantine}
\begin{tabular}{lcccc}
\toprule
\textbf{Byzantine \%} & \textbf{FedAvg} & \textbf{Krum} & \textbf{TrimMean} & \textbf{\parsfl} \\
\midrule
0\% & 89.4 & 88.2 & 88.8 & 89.1 \\
5\% & 74.2 & 86.1 & 87.4 & 88.3 \\
10\% & 52.3 & 83.7 & 86.1 & 87.6 \\
15\% & 34.1 & 78.4 & 83.8 & 86.2 \\
20\% & 21.7 & 71.2 & 80.3 & 84.1 \\
25\% & 18.3 & 52.4 & 64.7 & 68.3 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Communication Efficiency}

\begin{table}[H]
\centering
\caption{Communication Cost per Round (BERT-base, 110M params)}
\label{tab:comm}
\begin{tabular}{lccc}
\toprule
\textbf{Method} & \textbf{Upload (MB)} & \textbf{Rounds to 85\%} & \textbf{Total (GB)} \\
\midrule
FedAvg (fp32) & 440 & 50 & 22.0 \\
FedAvg (fp16) & 220 & 50 & 11.0 \\
SignSGD & 13.8 & 120 & 1.66 \\
Top-1\% + 8-bit & 3.5 & 80 & 0.28 \\
\parsfl adaptive & 2.2 & 65 & \textbf{0.14} \\
\bottomrule
\end{tabular}
\end{table}

\subsection{End-to-End Performance}

\begin{table}[H]
\centering
\caption{End-to-End Comparison (all protections enabled)}
\label{tab:e2e}
\begin{tabular}{lccccc}
\toprule
\textbf{System} & \textbf{SecAgg} & \textbf{DP} & \textbf{Byzantine} & \textbf{Compress} & \textbf{Acc.} \\
\midrule
Centralized (oracle) & -- & -- & -- & -- & 92.0 \\
FedAvg baseline & No & No & No & No & 89.4 \\
+ SecAgg & Yes & No & No & No & 89.4 \\
+ DP ($\varepsilon$=2) & Yes & Yes & No & No & 86.3 \\
+ Byzantine (20\%) & Yes & Yes & Yes & No & 84.1 \\
\parsfl (full) & Yes & Yes & Yes & Yes & 83.8 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Comparison with Existing FL Frameworks}

\begin{table}[H]
\centering
\caption{Feature Comparison with FL Frameworks}
\label{tab:frameworks}
\begin{tabularx}{\textwidth}{lXXXXX}
\toprule
\textbf{Feature} & \textbf{\parsfl} & \textbf{Flower} & \textbf{PySyft} & \textbf{TFF} & \textbf{FedML} \\
\midrule
Secure Aggregation & Threshold MPC & Plugin & SMPC & No & Plugin \\
Differential Privacy & RDP + adaptive & Plugin & Yes & Yes & Yes \\
Byzantine Resilience & Trimmed mean & No & No & No & Yes \\
Compression & Adaptive 3-stage & No & No & Partial & Yes \\
Decentralized & Committee & No & Partial & No & Yes \\
Censorship-resistant & Yes & No & No & No & No \\
Error feedback & Yes & No & No & No & No \\
\bottomrule
\end{tabularx}
\end{table}

\section{Deployment on Pars Network}

\subsection{Client Selection}

We use importance sampling for client selection, prioritizing clients with:

\begin{equation}
    P(\text{select client } i) \propto \frac{|D_i|}{\sum_j |D_j|} \cdot \mathsf{reliability}(i) \cdot \mathsf{bandwidth}(i)^{0.3}
\end{equation}

where $\mathsf{reliability}(i)$ is the historical completion rate of client $i$ and $\mathsf{bandwidth}(i)$ is the measured available bandwidth.

\subsection{Incentive Mechanism}

Clients are compensated in ASHA tokens proportional to their contribution:

\begin{equation}
    \text{reward}_i = \alpha_R \cdot |D_i| \cdot \mathsf{quality}(D_i) \cdot \mathsf{completion}(i)
\end{equation}

where $\mathsf{quality}(D_i)$ is estimated from the model improvement attributable to client $i$'s update (measured via Shapley value approximation), and $\mathsf{completion}(i) \in \{0, 1\}$.

\section{Limitations and Future Work}

\begin{enumerate}
    \item \textbf{Scalability:} The current secure aggregation protocol has $O(n^2)$ key agreement cost. Tree-based approaches could reduce this to $O(n \log n)$.
    \item \textbf{Heterogeneous Compute:} Clients with different device capabilities train at different speeds. Asynchronous FL protocols could improve utilization.
    \item \textbf{Verifiable Computation:} We assume semi-honest aggregators. Adding zero-knowledge proofs of correct aggregation would strengthen the trust model.
    \item \textbf{Personalization:} The current protocol trains a single global model. Local adaptation layers (e.g., LoRA) would improve per-client utility.
    \item \textbf{Foundation Models:} Extending to federated fine-tuning of large foundation models (7B+ parameters) requires new compression techniques.
\end{enumerate}

\section{Conclusion}

\parsfl provides a comprehensive federated learning framework tailored to the unique challenges of diaspora-scale collaborative AI training. By combining secure aggregation, differential privacy, Byzantine resilience, and adaptive compression, \parsfl enables the Pars Network to train community AI models without compromising individual privacy. Our theoretical analysis and empirical evaluation demonstrate that the framework achieves near-centralized accuracy while providing strong formal privacy guarantees, even in the presence of adversarial participants and unreliable communication.

\bibliographystyle{plainnat}

\begin{thebibliography}{35}

\bibitem[Abadi et~al.(2016)]{abadi2016deep}
Abadi, M., Chu, A., Goodfellow, I., et~al.
\newblock Deep learning with differential privacy.
\newblock In \emph{CCS}, pages 308--318, 2016.

\bibitem[Andrew et~al.(2021)]{andrew2021differentially}
Andrew, G., Thakkar, O., McMahan, H.~B., and Ramaswamy, S.
\newblock Differentially private learning with adaptive clipping.
\newblock In \emph{NeurIPS}, 2021.

\bibitem[Balle et~al.(2020)]{balle2020hypothesis}
Balle, B., Barthe, G., and Gavin, M.
\newblock Hypothesis testing interpretations and Renyi differential privacy.
\newblock In \emph{AISTATS}, pages 2496--2506, 2020.

\bibitem[Blanchard et~al.(2017)]{blanchard2017machine}
Blanchard, P., El~Mhamdi, E.~M., Guerraoui, R., and Stainer, J.
\newblock Machine learning with adversaries: Byzantine tolerant gradient descent.
\newblock In \emph{NeurIPS}, pages 118--128, 2017.

\bibitem[Bonawitz et~al.(2017)]{bonawitz2017practical}
Bonawitz, K., Ivanov, V., Kreuter, B., et~al.
\newblock Practical secure aggregation for privacy-preserving machine learning.
\newblock In \emph{CCS}, pages 1175--1191, 2017.

\bibitem[Dwork et~al.(2006)]{dwork2006calibrating}
Dwork, C., McSherry, F., Nissim, K., and Smith, A.
\newblock Calibrating noise to sensitivity in private data analysis.
\newblock In \emph{TCC}, pages 265--284, 2006.

\bibitem[Kairouz et~al.(2021)]{kairouz2021advances}
Kairouz, P., McMahan, H.~B., Avent, B., et~al.
\newblock Advances and open problems in federated learning.
\newblock \emph{Foundations and Trends in Machine Learning}, 14(1--2):1--210, 2021.

\bibitem[Li et~al.(2020)]{li2020federated}
Li, T., Sahu, A.~K., Zaheer, M., et~al.
\newblock Federated optimization in heterogeneous networks.
\newblock In \emph{MLSys}, 2020.

\bibitem[McMahan et~al.(2017)]{mcmahan2017communication}
McMahan, H.~B., Moore, E., Ramage, D., Hampson, S., and {Arcas}, B.~A.~y.
\newblock Communication-efficient learning of deep networks from decentralized data.
\newblock In \emph{AISTATS}, pages 1273--1282, 2017.

\bibitem[Mironov(2017)]{mironov2017renyi}
Mironov, I.
\newblock R\'enyi differential privacy.
\newblock In \emph{CSF}, pages 263--275, 2017.

\bibitem[Stich et~al.(2018)]{stich2018sparsified}
Stich, S.~U., Cordonnier, J.-B., and Jaggi, M.
\newblock Sparsified {SGD} with memory.
\newblock In \emph{NeurIPS}, pages 4447--4458, 2018.

\bibitem[Yin et~al.(2018)]{yin2018byzantine}
Yin, D., Chen, Y., Kannan, R., and Bartlett, P.
\newblock Byzantine-robust distributed learning: Towards optimal statistical rates.
\newblock In \emph{ICML}, pages 5650--5659, 2018.

\bibitem[Konecny et~al.(2016)]{konecny2016federated}
Kone\v{c}n\'y, J., McMahan, H.~B., Yu, F.~X., et~al.
\newblock Federated learning: Strategies for improving communication efficiency.
\newblock \emph{arXiv preprint arXiv:1610.05492}, 2016.

\bibitem[Geyer et~al.(2017)]{geyer2017differentially}
Geyer, R.~C., Klein, T., and Nabi, M.
\newblock Differentially private federated learning: A client level perspective.
\newblock \emph{arXiv preprint arXiv:1712.07557}, 2017.

\bibitem[Bell et~al.(2020)]{bell2020secure}
Bell, J.~H., Bonawitz, K.~A., Gasc\'on, A., Lepoint, T., and Raykova, M.
\newblock Secure single-server aggregation with (poly) logarithmic overhead.
\newblock In \emph{CCS}, pages 1253--1269, 2020.

\bibitem[Alistarh et~al.(2017)]{alistarh2017qsgd}
Alistarh, D., Grubic, D., Li, J., Tomioka, R., and Vojnovic, M.
\newblock {QSGD}: Communication-efficient {SGD} via gradient quantization and encoding.
\newblock In \emph{NeurIPS}, pages 1709--1720, 2017.

\bibitem[Wang et~al.(2021)]{wang2021field}
Wang, J., Charles, Z., Xu, Z., et~al.
\newblock A field guide to federated optimization.
\newblock \emph{arXiv preprint arXiv:2107.06917}, 2021.

\bibitem[Reddi et~al.(2021)]{reddi2021adaptive}
Reddi, S., Charles, Z., Zaheer, M., et~al.
\newblock Adaptive federated optimization.
\newblock In \emph{ICLR}, 2021.

\bibitem[Beutel et~al.(2020)]{beutel2020flower}
Beutel, D.~J., Tober, T., Steiner, R., et~al.
\newblock Flower: A friendly federated learning framework.
\newblock \emph{arXiv preprint arXiv:2007.14390}, 2020.

\bibitem[Ryffel et~al.(2018)]{ryffel2018pysyft}
Ryffel, T., Trask, A., Dahl, M., et~al.
\newblock A generic framework for privacy preserving deep learning.
\newblock \emph{arXiv preprint arXiv:1811.04017}, 2018.

\bibitem[He et~al.(2020)]{he2020fedml}
He, C., Li, S., So, J., et~al.
\newblock {FedML}: A research library for federated machine learning.
\newblock \emph{arXiv preprint arXiv:2007.13518}, 2020.

\bibitem[Yurochkin et~al.(2019)]{yurochkin2019bayesian}
Yurochkin, M., Agarwal, M., Ghosh, S., et~al.
\newblock Bayesian nonparametric federated learning of neural networks.
\newblock In \emph{ICML}, pages 7252--7261, 2019.

\bibitem[Ghosh et~al.(2020)]{ghosh2020efficient}
Ghosh, A., Chung, J., Yin, D., and Ramchandran, K.
\newblock An efficient framework for clustered federated learning.
\newblock In \emph{NeurIPS}, 2020.

\bibitem[Hu et~al.(2022)]{hu2022lora}
Hu, E.~J., Shen, Y., Wallis, P., et~al.
\newblock {LoRA}: Low-rank adaptation of large language models.
\newblock In \emph{ICLR}, 2022.

\bibitem[Shamir(1979)]{shamir1979share}
Shamir, A.
\newblock How to share a secret.
\newblock \emph{Communications of the ACM}, 22(11):612--613, 1979.

\bibitem[Dwork and Roth(2014)]{dwork2014algorithmic}
Dwork, C. and Roth, A.
\newblock The algorithmic foundations of differential privacy.
\newblock \emph{Foundations and Trends in Theoretical Computer Science}, 9(3--4):211--407, 2014.

\end{thebibliography}

\appendix

\section{Convergence Proof}
\label{app:proof}

We provide the full convergence proof for Theorem~\ref{thm:convergence}.

\begin{proof}
Starting from $L$-smoothness of $F$:
\begin{equation}
    F(w^{t+1}) \leq F(w^t) + \langle \nabla F(w^t), w^{t+1} - w^t \rangle + \frac{L}{2} \|w^{t+1} - w^t\|^2
\end{equation}

Let $\bar{\Delta}^t$ be the aggregated update after trimmed mean and decompression. Then $w^{t+1} = w^t - \eta \bar{\Delta}^t$.

We decompose the error:
\begin{align}
    \bar{\Delta}^t &= \underbrace{\nabla F(w^t)}_{\text{true gradient}} + \underbrace{(\bar{g}^t - \nabla F(w^t))}_{\text{stochastic noise}} + \underbrace{(\bar{g}_{\text{trim}}^t - \bar{g}^t)}_{\text{Byzantine error}} \\
    &\quad + \underbrace{(\hat{g}_{\text{trim}}^t - \bar{g}_{\text{trim}}^t)}_{\text{compression error}} + \underbrace{z^t}_{\text{DP noise}}
\end{align}

Taking expectations and applying standard bounds for each error term:

\begin{enumerate}
    \item Stochastic noise: $\EE[\|\bar{g}^t - \nabla F(w^t)\|^2] \leq \frac{\sigma_L^2}{n} + \tau^2 L^2 \sigma_G^2$
    \item Byzantine error: $\EE[\|\bar{g}_{\text{trim}}^t - \bar{g}^t\|^2] \leq O(f^2 G^2 / n^2)$ by~\citep{yin2018byzantine}
    \item Compression with error feedback: $\sum_t \EE[\|\hat{g}^t - \bar{g}^t\|^2] \leq \frac{1-\delta_c}{\delta_c} \sum_t \EE[\|\bar{g}^t\|^2]$
    \item DP noise: $\EE[\|z^t\|^2] = d \sigma^2 C^2 / n$
\end{enumerate}

Telescoping over $T$ rounds and dividing by $T$ yields the stated bound.
\end{proof}

\end{document}
