\documentclass[11pt,a4paper]{article}
\usepackage{../shared/parscover}

% ============================================================================
% Pars 2.0 — GPU-Native Sovereign L1 with DEX+EVM+FHE+SessionVM 4-Chain Set
% Launch retrospective: 2026-02-14
% ============================================================================

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{tikz}
\usepackage{caption}
\usepackage{fancyhdr}
\usepackage{abstract}
\usepackage{geometry}

\geometry{margin=1in}

\definecolor{parsred}{RGB}{204,0,0}
\definecolor{parsgold}{RGB}{218,165,32}

\hypersetup{
    colorlinks=true,
    linkcolor=parsred,
    citecolor=parsred,
    urlcolor=parsred
}

\lstset{
    basicstyle=\ttfamily\small,
    breaklines=true,
    frame=single,
    backgroundcolor=\color{gray!10},
    keywordstyle=\bfseries,
    commentstyle=\itshape,
    numbers=left,
    numberstyle=\tiny\color{gray},
    numbersep=5pt
}

\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{property}{Property}[section]
\newtheorem{invariant}{Invariant}[section]

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small Pars 2.0 Launch}
\fancyhead[R]{\small v2.0 --- 2026-02-14}
\fancyfoot[C]{\thepage}

\title{%
    \textbf{Pars 2.0 --- GPU-Native Sovereign L1 with DEX+EVM+FHE+SessionVM 4-Chain Set}\\[0.5em]
    \large From private-messaging substrate (1.0) to financially-complete sovereign L1\\
    for the Persian diaspora\\[0.4em]
    \normalsize Pars Network Launch Retrospective --- February 14, 2026
}

\author{%
    Zach Kelling\\
    Pars Network Foundation\\
    \texttt{research@pars.network}\\[0.4em]
    \small Version 2.0 --- Block-aligned activation 2026-02-14T00:00:00Z
}

\date{}

\begin{document}
\parscoverpage

\begin{abstract}
On 14 February 2026 the Pars Network activated its 2.0 epoch: a sovereign,
GPU-native Layer~1 carrying the canonical financial-rail core of every real
L1 in the Lux family --- a native decentralized exchange (DEX), an
Ethereum Virtual Machine execution layer (EVM), and a fully-homomorphic
encryption substrate (FHE). Pars 1.0 (2025) shipped the messaging,
identity, and federated-learning foundation that the diaspora needed but
omitted the financial rails; Pars 2.0 closes that gap without rewriting
1.0 primitives. The 2.0 chain runs Quasar consensus 4.0 with QuasarSTM
4.0 execution, enforces the GPU-Residency invariant (LP-137), and
participates in the cross-chain Liquidity Protocol launching 20 April
2026. This paper documents the 1.0$\to$2.0 transition, the financial-rail core
integration, the activation receipts captured at block height
$h_0 = 14{,}000{,}000$ on the Pars main DAG, and the production
parameters now in force.
\end{abstract}

\noindent\textbf{Keywords:} Pars Network, sovereign L1, GPU-native consensus,
DEX+EVM+FHE foundation, Persian diaspora, Liquidity Protocol, post-quantum.

\tableofcontents
\newpage

% ============================================================================
\section{Launch Summary}
\label{sec:launch}
% ============================================================================

\begin{table}[H]
\centering
\small
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Field} & \textbf{Value} \\
\midrule
Network name & Pars Network 2.0 \\
Activation block & $h_0 = 14{,}000{,}000$ on Pars main DAG \\
Activation timestamp & 2026-02-14T00:00:00Z (UTC, mainnet) \\
Chain ID (C-Chain) & $7070$ (mainnet), $7071$ (testnet) \\
Consensus & Quasar 4.0 (cf.\ LP-105) \\
Execution & QuasarSTM 4.0 (cf.\ LP-010 v4) \\
GPU residency & LP-137 enforced \\
EVM compatibility & Cancun + Pars precompiles 0x2000--0x2300, 0x0700 \\
FHE substrate & CKKS / TFHE hybrid (cf.\ \texttt{pars-fhe-privacy}) \\
DEX activation & 2026-02-14 (genesis pools) \\
Liquidity Protocol & Onboarding wave 2026-04-20 \\
Native token & PARS, fixed supply $10^9$ (no change from 1.0) \\
Reserve token & ASHA (cf.\ \texttt{pars-asha-reserve}) \\
Staking lock token & VeASHA (cf.\ \texttt{pars-veasha-tokenomics}) \\
\bottomrule
\end{tabularx}
\caption{Pars 2.0 launch parameters in force at $h_0$.}
\end{table}

The 2.0 launch is \emph{strictly additive}. No 1.0 primitive was removed.
The Session protocol (\texttt{pars-session-protocol}), mesh transport
(\texttt{pars-mesh-networking}), Persian-NLP precompile
(\texttt{pars-persian-nlp}), federated-learning protocol
(\texttt{pars-federated-learning}), and post-quantum suite
(\texttt{pars-post-quantum}) all remain operational with identical
on-wire formats. What changed is the addition of three first-class
financial subsystems on a GPU-native consensus engine.

% ============================================================================
\section{Pars 1.0 $\to$ 2.0 Evolution}
\label{sec:evolution}
% ============================================================================

\subsection{What 1.0 Shipped (2025)}

Pars 1.0 was a \emph{dual-layer} architecture (\texttt{pars-dual-layer}):
\begin{itemize}[leftmargin=*]
\item An EVM-compatible rollup posting state roots to Lux for security
inheritance, capable of $2{,}400$ TPS with 1.2-s confirmation.
\item A Session daemon layer providing end-to-end encrypted messaging with
forward secrecy and deniability, mesh-routable over LoRa/Bluetooth/IP.
\item A federated-learning fabric for Persian-language model training
without raw-data egress (\texttt{pars-federated-learning}).
\item A PQ migration plan committing to ML-DSA-87, ML-KEM-1024, and
SPHINCS+ for keys minted after activation
(\texttt{pars-post-quantum}).
\end{itemize}

The diaspora-sovereignty thesis was complete on the \emph{communication}
axis. The financial axis was deferred: 1.0 lacked a native DEX, lacked
private settlement, and lacked enough throughput in the EVM layer to
serve as a primary venue for tokenized assets.

\subsection{What 2.0 Adds (2026-02-14)}

Three first-class subsystems and one consensus upgrade:
\begin{enumerate}[leftmargin=*]
\item \textbf{Sovereign L1.} Pars graduates from rollup to standalone L1
running Quasar 4.0 consensus. State commitments to Lux remain available
as an \emph{optional} bridging mechanism but are no longer the security
root.
\item \textbf{Native DEX.} A GPU-native AMM (PARSWAP) with explicit
support for the Liquidity Protocol cross-chain settlement API, designed
for Persian-rial--denominated and IRT-pegged stablecoin flows.
\item \textbf{EVM 4.0.} Cancun-compatible EVM extended with Pars-specific
precompiles for calendar conversion (Jalali $\leftrightarrow$ Gregorian),
Pars Name Service (PNS), and Persian-NLP inference.
\item \textbf{FHE substrate.} CKKS / TFHE hybrid scheme exposed as a
precompile + sidecar, allowing confidential transactions, private
voting, and encrypted federated-learning aggregation
(\texttt{pars-fhe-privacy}).
\end{enumerate}

\subsection{Migration Path for 1.0 Users}

The 2.0 genesis state was constructed by the deterministic procedure of
Algorithm~\ref{alg:migrate}.

\begin{algorithm}[H]
\caption{Pars 1.0 $\to$ 2.0 state migration, applied at $h_0$}
\label{alg:migrate}
\begin{algorithmic}[1]
\Procedure{MigrateState}{$S_{1.0}, R_{1.0}$}
  \State $S_{2.0} \gets \emptyset$
  \For{each account $a \in S_{1.0}$}
    \State $S_{2.0}[a] \gets S_{1.0}[a]$ \Comment{balance, nonce, code, storage}
  \EndFor
  \State $S_{2.0}.\textsc{ChainID} \gets 7070$ \Comment{Pars C-Chain mainnet}
  \State $S_{2.0}.\textsc{ConsensusEpoch} \gets 4$
  \State $S_{2.0}.\textsc{ExecutionEpoch} \gets 4$
  \State $S_{2.0}.\textsc{GPUResidency} \gets \mathrm{enforced}$
  \State $S_{2.0}.\textsc{Precompiles}[\text{LXBook}\dots\text{LXFeed},\,\text{FHE}] \gets \mathrm{active}$
  \State $S_{2.0}.\textsc{FHESubstrate} \gets \mathrm{CKKS} \oplus \mathrm{TFHE}$
  \State $S_{2.0}.\textsc{LiquidityProtocol} \gets \mathrm{registered}$ \Comment{onboarding 2026-04-20}
  \State \Return $S_{2.0}$
\EndProcedure
\end{algorithmic}
\end{algorithm}

\begin{property}[Account preservation]
For every $a \in S_{1.0}$, the post-migration account
$S_{2.0}[a]$ has identical balance, nonce, deployed code, and storage
trie root as $S_{1.0}[a]$.
\end{property}

\begin{property}[Identity preservation]
Every Pars DID issued under \texttt{pars-did-standard} v1.0 verifies
unchanged against the 2.0 PDS resolver.
\end{property}

% ============================================================================
\section{GPU-Native Quasar 4.0 + QuasarSTM 4.0}
\label{sec:gpu}
% ============================================================================

\subsection{Why GPU-Native}

Pars 1.0 used CPU-bound execution. The throughput ceiling was set by
single-thread EVM interpretation and ECDSA verification. Pars 2.0
inherits the GPU-Residency invariant from Lux 4.0 (LP-137):

\begin{invariant}[GPU residency]
For every transaction $\tau$ executed at height $h$, the validator's
canonical execution path runs on a GPU device $g$ with all input
state, intermediate state, and output state resident on $g$'s HBM for
the duration of $\tau$. Any movement of state between host RAM and
device HBM during $\tau$ violates the invariant.
\end{invariant}

The invariant is enforced by QuasarSTM 4.0 transactional memory
boundaries (cf.\ LP-010 v4) and is observable via the
\texttt{gpu\_residency\_violations} Prometheus counter exported by every
Pars validator.

\subsection{Throughput at $h_0$}

\begin{table}[H]
\centering
\small
\begin{tabular}{lrrr}
\toprule
\textbf{Metric} & \textbf{Pars 1.0 (2025)} & \textbf{Pars 2.0 (2026-02-14)} & \textbf{Ratio} \\
\midrule
Sustained TPS (EVM) & $2{,}400$ & $48{,}000$ & $20\times$ \\
Confirmation latency & $1.2$ s & $180$ ms & $6.7\times$ \\
ECDSA verifications/s & $11{,}000$ & $620{,}000$ & $56\times$ \\
ML-DSA-87 verifications/s & --- & $84{,}000$ & --- \\
Block size cap & $30$M gas & $250$M gas & $8.3\times$ \\
\bottomrule
\end{tabular}
\caption{Throughput metrics observed on Pars mainnet validators at $h_0 + 1{,}000$.}
\end{table}

These numbers are within $4\%$ of the figures projected in
\texttt{pars-gpu-scaling} (2025).

% ============================================================================
\section{Pars 4-Chain Set: DEX + EVM + FHE + SessionVM}
\label{sec:chain-set}
% ============================================================================

Pars 2.0 runs exactly four chains: DEX (PARSWAP, D-Chain template),
EVM (C-Chain template), FHE (F-Chain template), and **SessionVM**
(S-Chain --- Pars-original). The first three are the financial-rail
core that Hanzo 4.0 ships as its canonical \emph{triumvirate}.
S-Chain is Pars-original: a post-quantum secure messaging substrate
wire-compatible with the Session/Oxen ecosystem, upgraded to
ML-DSA-65 + ML-KEM-768 for validator-quality post-quantum keys.
Pars is "triumvirate-plus-SessionVM" --- four chains exactly, no
local clones of Lux primary infrastructure.

Pars does not run its own P-Chain or X-Chain. Validator staking,
UTXO-asset workflows, threshold-key ceremonies (Q-Chain), ZK
rollups (Z-Chain), attestation (A-Chain, AIVM), bridging
(B-Chain), and MPC ceremonies (M-Chain) are accessed
\emph{natively} via the Lux primary network --- same Quasar cert
lanes, same subject-binding, same finality, no wrapped assets, no
cross-chain hop tax.

\subsection{DEX --- PARSWAP and Liquidity Protocol Integration}

PARSWAP is a constant-function automated market maker exposed as a
system contract at \texttt{0x0000\dots PA75}. Its design is intentionally
narrow: it is the venue for diaspora remittance flows and tokenized
real-asset settlement. Three classes of pool are supported:

\begin{enumerate}[leftmargin=*]
\item \textbf{Stable pools} (StableSwap-style invariant, $A=200$) for
IRT-pegged stablecoins, USDC, USDT, USDP-pegged synthetic rial.
\item \textbf{Weighted pools} (Balancer-style $\prod_i x_i^{w_i}=k$) for
diversified tokenized-asset baskets.
\item \textbf{Concentrated-liquidity pools} (Uniswap v3-style ticks)
for high-volume PARS pairs.
\end{enumerate}

\paragraph{Liquidity Protocol bridge.} Per the Liquidity Protocol common
API (LP-310), every pool exposes a deterministic settlement endpoint
that the cross-chain router can call to atomically rebalance liquidity
across Pars, Hanzo, Zoo, and Lux DEXes. The bridge wave activates
2026-04-20; the on-chain registration was committed at $h_0$.

The pool-state update equation for a stable pool is the StableSwap
invariant:
\begin{equation}
A \cdot n^n \cdot \sum_{i=1}^{n} x_i + D = A \cdot D \cdot n^n + \frac{D^{n+1}}{n^n \prod_{i=1}^{n} x_i}.
\label{eq:stableswap}
\end{equation}
Equation~\ref{eq:stableswap} is solved on-GPU per swap; the Newton
iteration converges in $\le 6$ rounds for all production parameters.

\subsection{EVM --- Pars Precompiles}

The Pars EVM is Cancun-equivalent and reuses the canonical Lux precompile
ranges (cf.\ Lux registry v0.4.x): \texttt{0x0600}--\texttt{0x06FF} for
post-quantum, \texttt{0x0700}--\texttt{0x07FF} for FHE/privacy,
\texttt{0x2000}--\texttt{0x23FF} for native DEX/HFT,
\texttt{0x1000}--\texttt{0x12FF} for X/T/Z-chain bridges. Pars adds three
precompiles at the empty \texttt{0x0c00} slot for diaspora-specific
functions:

\begin{table}[H]
\centering
\small
\begin{tabular}{lll}
\toprule
\textbf{Address} & \textbf{Function} & \textbf{Gas} \\
\midrule
\texttt{0x2000} & \textsc{LXBook} (HFT orderbook --- canonical Lux DEX) & $50{,}000$ \\
\texttt{0x2100} & \textsc{LXPool} (AMM pools --- canonical Lux DEX) & $50{,}000$ \\
\texttt{0x2300} & \textsc{LXFeed} (price feeds --- canonical Lux DEX) & $5{,}000$ \\
\texttt{0x0700} & \textsc{FHE} (CKKS / TFHE --- canonical Lux privacy) & $500{,}000$ \\
\texttt{0x0c01} & \textsc{CalendarConvert} (Jalali $\leftrightarrow$ Gregorian) & $250$ \\
\texttt{0x0c02} & \textsc{PNS\_Resolve} (Pars Name Service) & $1{,}500$ \\
\texttt{0x0c03} & \textsc{PersianNLP\_Tokenize} & $5{,}000 + 50/\text{token}$ \\
\bottomrule
\end{tabular}
\caption{Canonical and Pars-specific precompiles active from $h_0$.}
\end{table}

The PNS precompile makes domain resolution a constant-time on-chain
operation. The Persian-NLP precompile wraps the inference work
documented in \texttt{pars-persian-nlp} and \texttt{pars-distributed-inference}.
PARSWAP is a thin wrapper over LXPool plus a Liquidity-Protocol routing
adapter; it is not a fork.

\subsection{FHE --- Confidential Transactions and Private Voting}

Pars 2.0 ships the FHE substrate documented in \texttt{pars-fhe-privacy}
as a first-class L1 capability rather than a sidecar. Three application
patterns are activated at $h_0$:

\begin{enumerate}[leftmargin=*]
\item \textbf{Confidential transfers.} Amount-hiding transfers using
CKKS-encrypted amounts and a homomorphic balance-update circuit. Sender
and recipient are public; amounts are confidential.
\item \textbf{Private voting.} Per \texttt{pip-0104-private-voting-protocol},
ballots are TFHE-encrypted; tallying is performed under encryption; only
the final tally is decrypted by a threshold committee.
\item \textbf{Encrypted federated-learning aggregation.} Replaces the
1.0 secure-aggregation step of \texttt{pars-federated-learning} with
homomorphic gradient summation, removing the need for an honest
aggregator.
\end{enumerate}

The FHE precompiles operate on ciphertexts encoded under the threshold
public key $\mathsf{tpk}_{\text{Pars}}$ produced by a $t$-of-$n$
distributed key generation among Pars validators. Decryption requires
the cooperation of any $t$ validators; no single validator holds the
secret key in clear form.

% ============================================================================
\section{Tokenomics --- Preserved and Extended}
\label{sec:tokenomics}
% ============================================================================

The 1.0 tokenomics defined in \texttt{pars-tokenomics-staking},
\texttt{pars-asha-reserve}, and \texttt{pars-veasha-tokenomics} are
preserved unchanged. Three extensions accompany 2.0:

\begin{itemize}[leftmargin=*]
\item \textbf{DEX fee routing.} $0.30\%$ of swap volume routes to a
fee-distribution contract; the split is $50\%$ liquidity providers,
$30\%$ Asha reserve, $15\%$ VeAsha lockers, $5\%$ DAO treasury. This
mirrors the canonical Lux-family routing.
\item \textbf{FHE relay rewards.} A new emission stream, capped at $0.5\%$
of annual PARS supply, rewards validators that perform FHE bootstrapping
and ciphertext maintenance. Caps are enforced at the protocol level.
\item \textbf{Liquidity Protocol incentive vault.} A 2-year vault of
$2 \times 10^7$ PARS bootstraps cross-chain liquidity in the April 20
wave. Drawdown is gated by snapshot proofs of cross-chain volume.
\end{itemize}

\begin{theorem}[Supply preservation]
The 2.0 tokenomics changes preserve the fixed total supply of
$10^9$ PARS established in 1.0. The FHE relay rewards and Liquidity
Protocol incentive vault are funded from the previously-allocated
ecosystem-incentives bucket, not from new emission.
\end{theorem}
\begin{proof}
By construction. The 1.0 tokenomics paper allocates $30\%$ of total
supply ($3 \times 10^8$) to ecosystem incentives. The FHE relay reward
cap is $0.5\%$ of annual supply for $\le 8$ years $= 0.5\% \times 8 \times
10^9 = 4 \times 10^7$. The Liquidity Protocol incentive vault is
$2 \times 10^7$. Both fit within $3 \times 10^8$.
\end{proof}

% ============================================================================
\section{Cross-Chain Bridges}
\label{sec:bridges}
% ============================================================================

Pars 2.0 maintains canonical bridges to the three other Lux-family
networks. Bridge security relies on threshold-signed light-client proofs
plus a fallback fraud-proof window of 7 days.

\begin{table}[H]
\centering
\small
\begin{tabular}{lllr}
\toprule
\textbf{Counterparty} & \textbf{Mechanism} & \textbf{Proof} & \textbf{Settlement} \\
\midrule
Lux & A-Chain attestation (cf.\ \texttt{lux-achain-attestation}) & ML-DSA-87 & 24 s \\
Hanzo & Liquidity Protocol router & ML-DSA-87 & 30 s \\
Zoo & Liquidity Protocol router & ML-DSA-87 & 30 s \\
Ethereum & Optimistic light client & ECDSA + ML-DSA-87 & 7 d \\
\bottomrule
\end{tabular}
\caption{Pars 2.0 cross-chain bridges. All paths post-quantum at the signature layer.}
\end{table}

The Liquidity Protocol router unifies Pars--Hanzo and Pars--Zoo flows
under a single message format; the Lux bridge retains its native
attestation path because Lux is the root of trust for the family.

% ============================================================================
\section{Diaspora Tooling under Sanctions-Resilient Mesh}
\label{sec:diaspora}
% ============================================================================

Pars 2.0 keeps the diaspora as the primary user. Three new application
patterns are unlocked by the financial-rail core plus Persian extensions.

\subsection{Confidential Remittance}

A diaspora user in Toronto sends value to a recipient in Tehran. The
flow:
\begin{enumerate}[leftmargin=*]
\item Sender mints a CKKS-encrypted transfer to the recipient's stealth
address (\texttt{pars-stealth-addresses}).
\item Settlement occurs on Pars 2.0 with hidden amount.
\item Recipient claims via threshold decryption involving a Pars-Iran
relay node and at least two diaspora-resident validators.
\item Routing over the Pars mesh (\texttt{pars-mesh-networking})
provides delivery even when the recipient is under selective IP
blocking.
\end{enumerate}

\subsection{Tokenized Real-Asset DEX}

Real-asset issuance (gold receipts, commercial paper, fractional
real-estate units) on the Pars EVM is paired with PARSWAP pools.
Liquidity Protocol's cross-chain router exposes the same assets to the
Lux, Hanzo, and Zoo communities, multiplying liquidity without
replicating issuance.

\subsection{Private DAO Voting at Scale}

Pars DAO governance (\texttt{pars-dao-governance}) moves from clear-text
ballots to TFHE-encrypted voting. Ballots are submitted, tallied
homomorphically, and the result revealed; individual votes never leave
the encrypted domain. Coercion-resistance properties from
\texttt{pars-coercion-resistance} are preserved and strengthened.

% ============================================================================
\section{Activation Receipts}
\label{sec:receipts}
% ============================================================================

The launch was block-aligned at $h_0 = 14{,}000{,}000$. The receipts
below were captured by the Pars block explorer between $h_0$ and
$h_0 + 5{,}000$.

\begin{table}[H]
\centering
\small
\begin{tabular}{lll}
\toprule
\textbf{Event} & \textbf{Height} & \textbf{Notes} \\
\midrule
Quasar 4.0 epoch boundary & $h_0$ & cert lane id $0\text{x04}$ \\
QuasarSTM 4.0 activation & $h_0$ & STM root rotated \\
GPU-Residency invariant on & $h_0 + 1$ & first violation counter $= 0$ \\
PARSWAP genesis pools & $h_0 + 12$ & 7 stable, 4 weighted, 9 concentrated \\
FHE threshold key DKG & $h_0 + 24$ & $t = 67$, $n = 100$ \\
PNS root commitment & $h_0 + 30$ & Merkle root $0\text{x4f}\dots$ \\
Liquidity Protocol register & $h_0 + 48$ & router id $\mathrm{lp.pars.router}$ \\
First confidential transfer & $h_0 + 91$ & tx $0\text{x21}\dots$ \\
First private vote tally & $h_0 + 720$ & ballots $= 4{,}312$ \\
\bottomrule
\end{tabular}
\caption{Selected mainnet receipts from Pars 2.0 activation.}
\end{table}

% ============================================================================
\section{Conclusion}
\label{sec:conclusion}
% ============================================================================

Pars 2.0 promotes the network from a private-messaging substrate to a
financially-complete sovereign L1 without abandoning the diaspora
sovereignty thesis. The DEX provides the venue, the EVM provides the
logic, the FHE substrate provides the confidentiality, the GPU-native
consensus provides the speed, and the Liquidity Protocol exposes Pars
liquidity to the rest of the Lux family. Every 1.0 primitive remains
operational and unchanged.

The April 20, 2026 Liquidity Protocol activation will close the final
gap: cross-chain remittance and tokenized-asset flows under a single
common API, with Pars as a first-class participant.

\bibliographystyle{plain}
\begin{thebibliography}{99}

\bibitem{pars1} Pars Team. \emph{Pars Network: Sovereign Infrastructure for the Persian Diaspora --- Version 1.0}. February 2025.

\bibitem{lp010} Lux Network. \emph{LP-010: QuasarSTM 4.0 --- GPU-Native Transactional Memory}. 2026.

\bibitem{lp105} Lux Network. \emph{LP-105: Quasar Consensus 4.0}. 2026.

\bibitem{lp137} Lux Network. \emph{LP-137: GPU-Residency Invariant for Lux-Family L1s}. 2026.

\bibitem{lp310} Lux Network. \emph{LP-310: Liquidity Protocol Common Settlement API}. 2026.

\bibitem{parsfhe} Pars Network. \emph{Fully Homomorphic Encryption for Privacy-Preserving Computation on Pars Network --- PNP-009}. 2026.

\bibitem{parsasha} Pars Network. \emph{Asha Reserve --- PNP-018}. 2026.

\bibitem{parsveasha} Pars Network. \emph{VeAsha Tokenomics --- PNP-019}. 2026.

\bibitem{parsdid} Pars Network. \emph{Pars DID Standard --- PNP-006}. 2026.

\bibitem{parsmesh} Pars Network. \emph{Pars Mesh Networking --- PNP-002}. 2026.

\bibitem{parspq} Pars Network. \emph{Pars Post-Quantum Suite --- PNP-007}. 2026.

\bibitem{parsfl} Pars Network. \emph{Pars Federated Learning --- PNP-014}. 2026.

\bibitem{parsstealth} Pars Network. \emph{Pars Stealth Addresses --- PNP-011}. 2026.

\bibitem{parscoer} Pars Network. \emph{Pars Coercion Resistance --- PNP-004}. 2026.

\bibitem{parsdao} Pars Network. \emph{Pars DAO Governance --- PNP-016}. 2026.

\bibitem{parsnlp} Pars Network. \emph{Pars Persian-NLP --- PNP-008}. 2026.

\bibitem{parsdistinf} Pars Network. \emph{Pars Distributed Inference --- PNP-015}. 2026.

\end{thebibliography}

\end{document}
